---
title: "Project1"
format: html
editor: visual
---

## Data Processing

### First Steps

In this part, I will read in and parse data step and step.

Read in data set

```{r}
library(tidyverse)
library(tidyr)
dat <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv", col_names = TRUE)
```

1.  Select columns: Area_name, STCOU, and any column ends in "D"

```{r}
selected_dat <- dat |> 
  select(Area_name, STCOU, ends_with("D")) |>
  rename("area_name" = "Area_name")
```

2.  Convert the data into long format

```{r}
long_dat <- selected_dat |>
  pivot_longer(cols = 3:12,
               names_to = "enrollment",
               values_to = "number")
```

3.  Parse enrollment into two new columns

```{r}
parse_dat <- long_dat |>
  separate_wider_position(enrollment, c(measurement = 3, 4, year = 2, 1))|>
  mutate(year = ifelse(as.numeric(year) < 10, as.numeric(year) + 2000, as.numeric(year) + 1900))
```

4.  create non-county data and county data

```{r}
county_dat <- parse_dat |>
  filter(grepl(pattern = ", \\w\\w", area_name)) 
class(county_dat) <- c("county", class(county_dat))

non_county_dat <- parse_dat |>
  filter(!grepl(pattern = ", \\w\\w", area_name)) 
class(non_county_dat) <- c("state", class(non_county_dat))
```

5.  Create new variable for county data

```{r}
final_county_dat <- county_dat |>
  mutate(state = substr(area_name, nchar(area_name)-1, nchar(area_name)))
```

6.  Add division information to non-county data

```{r}
final_non_county_dat <- non_county_dat |>
  mutate(division = case_when(
                        area_name %in% c("CONNECTICUT", "MAINE", "MASSACHUSETTS", "NEW HAMPSHIRE", "RHODE ISLAND", "VERMONT") ~ "Division 1",
                        area_name %in% c('NEW JERSEY', 'NEW YORK', 'PENNSYLVANIA') ~ "Division 2",
                        area_name %in% c('ILLINOIS', 'INDIANA', 'MICHIGAN', 'OHIO', 'WISCONSIN') ~ "Division 3",
                        area_name %in% c('IOWA', 'KANSAS', 'MINNESOTA', 'MISSOURI', 'NEBRASKA', 'NORTH DAKOTA', 'SOUTH DAKOTA') ~ "Division 4",
                        area_name %in% c('DELAWARE', 'FLORIDA', 'GEORGIA', 'MARYLAND', 'NORTH CAROLINA', 'SOUTH CAROLINA', 'VIRGINIA', 'DISTRICT OF COLUMBIA', 'District of Columbia', 'WEST VIRGINIA') ~ "Division 5",
                        area_name %in% c('ALABAMA', 'KENTUCKY', 'MISSISSIPPI', 'TENNESSEE') ~ "Division 6",
                        area_name %in% c('ARKANSAS', 'LOUISIANA', 'OKLAHOMA', 'TEXAS') ~ "Division 7",
                        area_name %in% c('ARIZONA', 'COLORADO', 'IDAHO', 'MONTANA', 'NEVADA', 'NEW MEXICO', 'UTAH', 'WYOMING') ~ "Division 8",
                        area_name %in% c('ALASKA', 'CALIFORNIA', 'HAWAII', 'OREGON', 'WASHINGTON') ~ "Division 9",
                        area_name == "UNITED STATES" ~ "ERROR",
 ) 
) 
```

### Then, I will write functions to do above pieces and call it all

Function does steps 1 & 2

```{r}
function_for_step_1_2 <- function(dat, default_var_name = "number") {
  result_from_step2 <- dat |> 
    select(Area_name, STCOU, ends_with("D")) |>
    rename("area_name" = "Area_name") |>
    pivot_longer(cols = 3:12,
               names_to = "enrollment",
               values_to = default_var_name)
}
```

Function for step 3

```{r}
function_for_step_3 <- function(result_from_step2) {
  result_from_step3 <- result_from_step2 |>
    separate_wider_position(enrollment, c(measurement = 3, 4, year = 2, 1))|>
    mutate(year = ifelse(as.numeric(year) < 10, as.numeric(year) + 2000, as.numeric(year) + 1900))
}
```

Function for step 5

```{r}
function_for_step_5 <- function(county_dat){
  county_dat |>
    mutate(state = substr(area_name, nchar(area_name)-1,
                          nchar(area_name)))
}
```

Function for step 6

```{r}
function_for_step_6 <- function(non_county_dat){
  non_county_dat |>
    mutate(division = case_when(
                        area_name %in% c("CONNECTICUT", "MAINE", "MASSACHUSETTS", "NEW HAMPSHIRE", "RHODE ISLAND", "VERMONT") ~ "Division 1",
                        area_name %in% c('NEW JERSEY', 'NEW YORK', 'PENNSYLVANIA') ~ "Division 2",
                        area_name %in% c('ILLINOIS', 'INDIANA', 'MICHIGAN', 'OHIO', 'WISCONSIN') ~ "Division 3",
                        area_name %in% c('IOWA', 'KANSAS', 'MINNESOTA', 'MISSOURI', 'NEBRASKA', 'NORTH DAKOTA', 'SOUTH DAKOTA') ~ "Division 4",
                        area_name %in% c('DELAWARE', 'FLORIDA', 'GEORGIA', 'MARYLAND', 'NORTH CAROLINA', 'SOUTH CAROLINA', 'VIRGINIA', 'DISTRICT OF COLUMBIA', 'District of Columbia', 'WEST VIRGINIA') ~ "Division 5",
                        area_name %in% c('ALABAMA', 'KENTUCKY', 'MISSISSIPPI', 'TENNESSEE') ~ "Division 6",
                        area_name %in% c('ARKANSAS', 'LOUISIANA', 'OKLAHOMA', 'TEXAS') ~ "Division 7",
                        area_name %in% c('ARIZONA', 'COLORADO', 'IDAHO', 'MONTANA', 'NEVADA', 'NEW MEXICO', 'UTAH', 'WYOMING') ~ "Division 8",
                        area_name %in% c('ALASKA', 'CALIFORNIA', 'HAWAII', 'OREGON', 'WASHINGTON') ~ "Division 9",
                        area_name == "UNITED STATES" ~ "ERROR",
 ) 
) 
}
```

Function for steps 4, 5, 6

```{r}
function_for_steps4_5_6 <- function(result_from_step3){
  county_dat <- result_from_step3 |>
    filter(grepl(pattern = ", \\w\\w", area_name)) |>
    function_for_step_5()
  class(county_dat) <- c("county", class(county_dat))
  
  non_county_dat <- result_from_step3 |>
    filter(!grepl(pattern = ", \\w\\w", area_name)) |>
    function_for_step_6()
  class(non_county_dat) <- c("state", class(non_county_dat))
  return(list(county_dat, non_county_dat))
}
```

Create a wrapper function

```{r}
my_wrapper <- function(url, default_var_name = "number"){
  result <- read_csv(url, col_names = TRUE) |>
    function_for_step_1_2(default_var_name = "number") |>
    function_for_step_3() |>
    function_for_steps4_5_6()
return(result)
}
```


### Then, we write a combine function to combine the reuslts from two data sets.

```{r}
my_combine <- function(x, y){
  county_dat <- bind_rows(x[1], y[1])
  non_county_dat <- bind_rows(x[2], y[2])
  return(list(county_dat, non_county_dat))
}
```

## Data Sumarizing

### Firstly, we write a function to plot state data.

```{r}
plot.state <- function(df, var_name = "number"){
  df$year <- as.factor(df$year)
  ggplot(data = df |>
           filter(division != "ERROR") |>
           group_by(year, division) |>
           summarize(Enrollment_mean = mean(get(var_name))), 
         aes(x = year, y = Enrollment_mean, group = division, color = division)) +
           geom_line() +
           geom_point() +
           labs(x = "Year")
}

```

### Then, we write a function to plot county data.

```{r}
plot.county <- function(df, Interest_state = "NC", var_name = "number", group = "Top", n = 5){
  df$year <- as.factor(df$year)
  
  ranking <- df |> 
    filter(state == Interest_state) |>
    group_by(area_name) |>
    summarize(county_mean = mean(get(var_name))) |>
    arrange(desc(county_mean))
      
  if (group == "Top") {
    selected_area_name <- ranking$area_name[1:n]
  } else if (group == "Bottom") {
    selected_area_name <- tail(ranking$area_name, n)
  }
    
  ggplot(data = df |>
           filter(area_name %in% selected_area_name) |>
           group_by(year, area_name) |>
           summarize(Enrollment_mean = mean(get(var_name))), 
         aes(x = year, y = Enrollment_mean, group = area_name, color = area_name)) +
           geom_line() +
           geom_point() +
           labs(x = "Year")
}

```

## Put it Together

### Firstly, we call the wrapper function twice to read in and parse two data sets.

```{r}
dat1 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv", default_var_name)

dat2 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv", default_var_name)
```

### Then, we call the combining function to put two data sets into one object

```{r}
Final_dat <- my_combine(x=dat1, y=dat2)
```

### Use the plot function on the state data frame.

```{r}
plot.state(df = Final_dat[[2]], var_name = "number")
```
### Use the plot function on the county data frame

Once specifying the state to be “NC”, the group being the top, the number looked at being 20

```{r}
plot.county(df=Final_dat[[1]], Interest_state = "NC", var_name = "number", group = "Top", n = 20)
```
Once specifying the state to be “SC”, the group being the bottom, the number looked at being 7

```{r}
plot.county(df=Final_dat[[1]], Interest_state = "SC", var_name = "number", group = "Bottom", n = 7)
```
Once without specifying anything (defaults used)

```{r}
plot.county(df=Final_dat[[1]])
```
Once specifying the state to be “PA”, the group being the top, the number looked at being 8

```{r}
plot.county(df=Final_dat[[1]], Interest_state = "PA", var_name = "number", group = "Top", n = 8)
```

### Lastly, read in another couple similar data sets and apply our functions!

```{r}
dat3 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01a.csv", default_var_name)

dat4 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01b.csv", default_var_name)

dat5 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01c.csv", default_var_name)

dat6 <- my_wrapper("https://www4.stat.ncsu.edu/~online/datasets/PST01d.csv", default_var_name)
```

### Then, we call the combining function to put above four data sets into one object

```{r}
dat3_4 <- my_combine(x=dat3, y=dat4)
dat3_4_5 <- my_combine(x=dat3_4, y=dat5)
Final_dat2 <- my_combine(x=dat3_4_5, y=dat6)
```

### Use the plot function on the state data frame.

```{r}
plot.state(df = Final_dat2[[2]], var_name = "number")
```

### Use the plot function on the county data frame

Once specifying the state to be “CA”, the group being the top, the number looked at being 15

```{r}
plot.county(df=Final_dat2[[1]], Interest_state = "CA", var_name = "number", group = "Top", n = 15)
```
Once specifying the state to be “TX”, the group being the top, the number looked at being 4

```{r}
plot.county(df=Final_dat2[[1]], Interest_state = "TX", var_name = "number", group = "Top", n = 4)
```
Once without specifying anything (defaults used)

```{r}
plot.county(df=Final_dat2[[1]])
```
Once specifying the state to be “NY”, the group being the top, the number looked at being 10

```{r}
plot.county(df=Final_dat2[[1]], Interest_state = "NY", var_name = "number", group = "Top", n = 10)
```

